Brief notes on erosion

### 2017-07-16 docs/TexNotes/Notes.tex
Writing Latex notes on permeability, resistance, and how to remove the contribution from the walls

\section{Permeability, conductivity, drag, resistance, and all of that}

Darcy's law is
\begin{equation}
\bq = - \bvec{K} \grad \phi
\end{equation}
where $\bq$ is the specific discharge, $\phi = z + p/\gamma$ is the hydraulic head, and $\gamma = \rho g$. The rank-2 tensor $\bvec{K}$ is the {\em conductivity}, which is related to the {\em permeability} $\bvec{k}$ via $\bvec{K} = \bvec{k} \gamma/\mu$, where $\mu$ is the viscosity of the fluid. The units of conductivity are length-per-time (i.e.~velocity) and the units of permeability are length squared.
Perhaps the most standard parameter to characterize the porous medium is the permeability, so lets focus on that. Let us also eliminate gravity from consideration to get
\begin{equation}
\bq = - \bvec{k} \grad p / \mu
\end{equation}
Consider first an isotropic medium with permeability $k$ (scalar). I propose that we decompose the permeability into the contribution from channel walls and from the embedded bodies $k = k_{c} + k_{b}$. Lets first figure out the contribution from the channel walls $k_c$. Consider flow in a channel of width $2w$ due to a pressure difference of $\Delta p$ occurring over a length $2 \ell$, so that the pressure gradient is $\grad p = -\Delta p/(2 \ell)$. If I did it correctly, the velocity is
\begin{equation}
u(z) = \frac{\Delta p}{4 \mu \ell} (w^2 - z^2)
\end{equation}
Then we can calculate the discharge as
\begin{equation}
q = \frac{1}{2w} \int_{-w}^{w} u(z) dz = \frac{\Delta p w^2}{6 \mu \ell} = - \frac{w^2}{3 \mu} \grad p
\end{equation}
so that the channel permeability is given by
\begin{equation}
k_c = w^2/3
\end{equation}
This result at least has the correct units (length squared).

>>>>

\section{Permeability, conductivity, drag, resistance, and all that}

Darcy's law is
\begin{equation}
\bq = - \bvec{K} \grad \phi
\end{equation}
where $\bq$ is the specific discharge, $\phi = z + p/\gamma$ is the hydraulic head, and $\gamma = \rho g$. The rank-2 tensor $\bvec{K}$ is the {\em conductivity}, which is related to the {\em permeability} $\bvec{k}$ via $\bvec{K} = \bvec{k} \gamma/\mu$, where $\mu$ is the viscosity of the fluid. The units of conductivity are length-per-time (i.e.~velocity) and the units of permeability are length squared.
Perhaps the most standard parameter to characterize the porous medium is the permeability, so lets focus on that. Let us also eliminate gravity from consideration to get
\begin{equation}
\bq = - \bvec{k} \grad p / \mu
\end{equation}
Consider first an isotropic medium with permeability $k$ (scalar). As a warmup, consider flow in a channel of width $2w$ due to a pressure difference of $\Delta p$ occurring over a length $2 \ell$, so that the pressure gradient is $\grad p = -\Delta p/(2 \ell)$. If I did it correctly, the velocity is
\begin{equation}
u(z) = \frac{\Delta p}{4 \mu \ell} (w^2 - z^2)
\end{equation}
Then we can calculate the discharge as
\begin{equation}
q = \frac{1}{2w} \int_{-w}^{w} u(z) dz = \frac{\Delta p w^2}{6 \mu \ell} = - \frac{w^2}{3 \mu} \grad p
\end{equation}
so that the permeability due to the channel walls is given by
\begin{equation}
k_w = w^2/3
\end{equation}
This result at least has the correct units (length squared).

Next, consider measuring the permeability for flow through a matrix of bodies. To accomplish this, we should measure the pressure and velocity along two vertical lines, one at $x = x_0$ and another at $x = -x_0$, and average to get $p^+ = \left< p(x_0) \right>$, $p^- = \left<p(-x_0) \right>$ and the same for $q$. Then the measured value of the total permeability (setting $\mu = 1$), is
\begin{equation}
k_{tot} = \frac{x_0 (q^+ + q^-)}{p^- - p^+}
\end{equation}
Presumably, $q$ (the average along a vertical cross-section) should be the same everywhere, but we should check this.

The `total' permeability above includes resistance from both the bodies and the channel walls, and so the value is artificially low (the true permeability of just the bodies will be a bit higher). In electrical circuits, the conductivity and resistivity are simply reciprocals of one another, so I presume the same holds for porous media flow, giving
\begin{equation}
\frac{1}{k_{tot}} = \frac{1}{k_w} + \frac{1}{k_b}
\end{equation}
 where $k_w = w^2/3$ is from the channel walls and $k_b$ is from the erodible bodies. Then, to compute the permeability from the bodies, we would remove the contribution from the walls. First, lets consider the resistances (the reciprocals of the permeabilities)
\begin{equation}
\frac{1}{k_b} = \frac{1}{k_{tot}} - \frac{1}{k_w} = \frac{k_w - k_{tot}}{k_w k_{tot}}
\end{equation}
which gives
\begin{equation}
k_b = \frac{k_w k_{tot}}{k_w - k_{tot}}
\end{equation}
